\LARGE \bf Stochastic Uncertainty Propagation in Power System Dynamics using Measure-valued Proximal Recursions

### 2021-07-05 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS.tex
updating Sec. V

\subsubsection{Notation and organization}
We use boldfaced small letters for vectors, and boldfaced capital letters for matrices. The set of natural numbers, real numbers and complex numbers are denoted as $\mathbb{N}$, $\mathbb{R}$, and $\mathbb{C}$, respectively. The symbol $\nabla_{\bm{x}}$ denotes the Euclidean gradient operator with respect to (w.r.t.) the vector $\bm{x}$. Thus, $\nabla_{\bm{x}}\cdot$ stands for the divergence, and $\Delta_{\bm{x}}$ stands for the Laplacian w.r.t. vector $\bm{x}$. We use $\langle\cdot,\cdot\rangle$ to denote the standard Euclidean inner product. The real and imaginary parts of a complex number $z$ are denoted via $\Re(z)$ and $\Im(z)$, respectively. We use the superscript $*$ to denote the complex conjugate. The uniform probability distribution over an interval $[a,b]$ is denoted as ${\rm{Unif}}\left([a,b]\right)$. Likewise, the $n$ dimensional uniform probability distribution over $[a,b]^{n}$ is denoted as ${\rm{Unif}}\left([a,b]^{n}\right)$. The symbol $\propto$ denotes proportionality, $|\cdot|$ denotes the absolute magnitude, $\|\cdot\|_{2}$ denotes the standard Euclidean 2-norm, $\otimes$ denotes the Kronecker product, $\odot$ and $\oslash$ respectively denote the elementwise (Hadamard) product and division, $\det(\cdot)$ stands for the determinant, and the subscript $_{\sharp}$ denotes the pushforward of a PDF via a map. The $n\times n$ identity and zero matrices are denoted as $\bm{I}_{n}$ and $\bm{0}_{n\times n}$, respectively.

@@

%\subsubsection{Deterministic dynamics and network reduced model}
%Before considering stochastic uncertainties in the dynamics, let us briefly recall the network reduced power systems model \cite[Ch. 7]{sauerpai1998} in the deterministic case. Suppose we denote the collection of generator buses in a networked power system as $\mathcal{G}$, and the collection of non-generator buses as $\mathcal{N}$. Assuming that the non-generator buses are modeled as constant current loads, with currents $I_i$, $i \in \mathcal{N}$, the power flow equations become
%\begin{subequations}
%\begin{eqnarray}
%m_i \ddot{\theta} + \gamma_i \dot{\theta_i} = P_{mech,i} - P_{load,i} - \sum_{j} |V_i| |V_j| |Y_{ij}| \sin(\theta_i - \theta_j + \phi_{ij}) &~ i \in \mathcal{G}~\text{buses with generators} \\
%0 = - \text{Real}(V_i I_i^\dagger) - \sum_{j} |V_i| |V_j| |Y_{ij}| \sin(\theta_i - \theta_j + \phi_{ij}) &~ i \in \mathcal{L}~\text{buses without generators} \\
%\end{eqnarray}
%\end{subequations}
%Note: $P_{mech,i} \in \mathbb{R}$, $P_{load,i} \in \mathbb{R}$ and $I_i \in \mathbb{C}$ are all give parameters; load current is a complex number. 
%
%
%\subsubsection{Stochastic dynamics}
In this work, we consider the coupled stochastic differential equations (SDEs) associated with the networked-reduced power systems model \cite[Ch. 7]{sauerpai1998}. Specifically, for a power network with $n$ generators, the stochastic dynamics for the $i$-th generator is given by the It\^{o} SDEs
\begin{subequations}
\begin{align}
&{\mathrm{d}}\theta_{i} = \omega_{i}\:{\mathrm{d}}t,\label{RotAngleGen}\\
&m_{i}\:{\mathrm{d}}\omega_{i} =	\left(\!P_{i} - \gamma_{i}\omega_{i} - \displaystyle\sum_{j=1}^{n} k_{ij}\sin\left(\theta_{i}-\theta_{j}-\varphi_{ij}\right)\!\right){\mathrm{d}}t \nonumber\\
&\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad+\sigma_{i}\:{\mathrm{d}}w_{i},
\end{align}
\label{ItoSDEcomponentlevel}
\end{subequations}
where the state variables are the rotor angles $\theta_{i}\in[0,2\pi)$ and the rotor angular velocities $\omega_{i}\in\mathbb{R}$, for $i\in\{1,\hdots,n\}$. The stochastic forcing is modeled through the standard Wiener process $w_{i}(t)$, and the diffusion coefficient $\sigma_{i}>0$ denotes the intensity of stochastic forcing at the $i$-th generator.

@@

With the $i$th generator, we associate its inertia $m_{i}>0$ and damping coefficient $\gamma_{i}>0$. The other parameters: the \emph{effective power input} $P_{i}$, the \emph{phase shift} $\varphi_{ij}\in[0,\frac{\pi}{2})$, and the \emph{coupling coefficients} $k_{ij}\geq 0$, depend on the network reduced admittance matrix $\bm{Y}\equiv[Y_{ij}]_{i,j=1}^{n}\in\mathbb{C}^{n\times n}$. Specifically,
\begin{subequations}
\begin{align}
 P_{i} &= P_{i}^{\text{mech}} - P_{i}^{\text{load}} - |V_{i}|^{2} \Re\left(Y_{ii}\right) + \Re\left(V_i \cdot I_{i}^{*}\right),\\
 \varphi_{ij} &= \begin{cases} -\arctan\left(\dfrac{\Re\left(Y_{ij}\right)}{\Im\left(Y_{ij}\right)}\right), & \text{if}\;i\neq j,\\
 0, & \text{otherwise},
 \end{cases}\\
 k_{ij} &= \begin{cases} |V_{i}| |V_{j}| \vert Y_{ij}\vert, & \text{if}\;i\neq j,\\
 0, & \text{otherwise},
 \end{cases}
\end{align}
\end{subequations}
where $P_{i}^{\text{mech}}$ is the mechanical power input, $P_{i}^{\text{load}}$ is {\red{the real load (ignoring reactive load)}}, and $|V_{i}|$ is the voltage magnitude for generator $i$. {\red{{\textbf{Xinbo:}} One sentence saying any approximation made in (2) by using $V_{i}$ instead of the internal voltage $E_{i}$}}

@@

\begin{remark}
If the unreduced network has $n$ generators and $m$ buses, then the unreduced admittance matrix $\bm{Y}_{\rm{unreduced}} \in \mathbb{C}^{(n+m)\times(n+m)}$ satisfies
\begin{align}
\bm{Y}_{\rm{unreduced}} = .
\label{unreducedY}	
\end{align}
The network reduced admittance matrix $\bm{Y}\in\mathbb{C}^{n\times n}$ is the Schur complement
\[\bm{Y} = \bm{Y}_{\rm{unreduced}}.\]	
\end{remark}

@@

\begin{figure}[htpb]
\centering
\includegraphics[width=0.9\linewidth]{ProxSchmatic.pdf}
\caption{\small{A schematic of approximating the flow of the joint PDF trajectory $\rho(t,\bm{x})$ via the sequence $\{\varrho_{k}\}_{k\in\mathbb{N}}$ generated by a variational recursion over $\mathcal{P}_{2}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right)$. The curve $\rho(t,\bm{x})$ solves a kinetic Fokker-Planck PDE initial value problem. While the joint PDFs are supported over the finite dimensional base manifold $\mathbb{T}^{n}\times\mathbb{R}^{n}$, the proximal updates $\{\varrho_{k}\}_{k\in\mathbb{N}}$ (shown as circled markers with no face-color) evolve over the infinite dimensional manifold $\mathcal{P}_{2}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right)$.}}
\vspace*{-0.1in}
\label{fig:ProxSchematic}
\end{figure}

>>>>

\subsubsection{Notation and organization}
%We use boldfaced small letters for vectors, and boldfaced capital letters for matrices. 
The set of natural numbers, real numbers and complex numbers are denoted as $\mathbb{N}$, $\mathbb{R}$, and $\mathbb{C}$, respectively. The symbol $\nabla_{\bm{x}}$ denotes the Euclidean gradient operator with respect to (w.r.t.) the vector $\bm{x}$. Thus, $\nabla_{\bm{x}}\cdot$ stands for the divergence, and $\Delta_{\bm{x}}$ stands for the Laplacian w.r.t. vector $\bm{x}$. We use $\langle\cdot,\cdot\rangle$ to denote the standard Euclidean inner product. The real and imaginary parts of a complex number $z$ are denoted via $\Re(z)$ and $\Im(z)$, respectively. We use the superscript $*$ to denote the complex conjugate. The uniform probability distribution over an interval $[a,b]$ is denoted as ${\rm{Unif}}\left([a,b]\right)$. Likewise, the $n$ dimensional uniform probability distribution over $[a,b]^{n}$ is denoted as ${\rm{Unif}}\left([a,b]^{n}\right)$. The symbol $\propto$ denotes proportionality, $|\cdot|$ denotes the absolute magnitude, $\|\cdot\|_{2}$ denotes the standard Euclidean 2-norm, $\otimes$ denotes the Kronecker product, $\odot$ and $\oslash$ respectively denote the elementwise (Hadamard) product and division, $\det(\cdot)$ stands for the determinant, and the subscript $_{\sharp}$ denotes the pushforward of a PDF via a map. The $n\times n$ identity and zero matrices are denoted as $\bm{I}_{n}$ and $\bm{0}_{n\times n}$, respectively.

@@

%\subsubsection{Deterministic dynamics and network reduced model}
%Before considering stochastic uncertainties in the dynamics, let us briefly recall the network reduced power systems model \cite[Ch. 7]{sauerpai1998} in the deterministic case. Suppose we denote the collection of generator buses in a networked power system as $\mathcal{G}$, and the collection of non-generator buses as $\mathcal{N}$. Assuming that the non-generator buses are modeled as constant current loads, with currents $I_i$, $i \in \mathcal{N}$, the power flow equations become
%\begin{subequations}
%\begin{eqnarray}
%m_i \ddot{\theta} + \gamma_i \dot{\theta_i} = P_{mech,i} - P_{load,i} - \sum_{j} |V_i| |V_j| |Y_{ij}| \sin(\theta_i - \theta_j + \phi_{ij}) &~ i \in \mathcal{G}~\text{buses with generators} \\
%0 = - \text{Real}(V_i I_i^\dagger) - \sum_{j} |V_i| |V_j| |Y_{ij}| \sin(\theta_i - \theta_j + \phi_{ij}) &~ i \in \mathcal{L}~\text{buses without generators} \\
%\end{eqnarray}
%\end{subequations}
%Note: $P_{mech,i} \in \mathbb{R}$, $P_{load,i} \in \mathbb{R}$ and $I_i \in \mathbb{C}$ are all give parameters; load current is a complex number. 
%
%
%\subsubsection{Stochastic dynamics}
In this work, we consider the coupled stochastic differential equations (SDEs) associated with the networked-reduced power systems model \cite[Ch. 7]{sauerpai1998}. Specifically, for a power network with $n$ generators, the stochastic dynamics for the $i$-th generator is given by the It\^{o} SDEs
\begin{subequations}
\begin{align}
&{\mathrm{d}}\theta_{i} = \omega_{i}\:{\mathrm{d}}t,\label{RotAngleGen}\\
&m_{i}\:{\mathrm{d}}\omega_{i} =	\left(\!P_{i} - \gamma_{i}\omega_{i} - \displaystyle\sum_{j=1}^{n} k_{ij}\sin\left(\theta_{i}-\theta_{j}-\varphi_{ij}\right)\!\right){\mathrm{d}}t \nonumber\\
&\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad+\sigma_{i}\:{\mathrm{d}}w_{i},
\end{align}
\label{ItoSDEcomponentlevel}
\end{subequations}
where the state variables are the rotor angles $\theta_{i}\in[0,2\pi)$ and the rotor angular velocities $\omega_{i}\in\mathbb{R}$, for $i\in\{1,\hdots,n\}$. The stochastic forcing is modeled through the standard Wiener process $w_{i}(t)$, and the diffusion coefficient $\sigma_{i}>0$ denotes the intensity of stochastic forcing at the $i$th generator.

@@

With the $i$th generator, we associate its inertia $m_{i}>0$ and damping coefficient $\gamma_{i}>0$. The other parameters: the \emph{effective power input} $P_{i}$, the \emph{phase shift} $\varphi_{ij}\in[0,\frac{\pi}{2})$, and the \emph{coupling coefficients} $k_{ij}\geq 0$, depend on the network reduced admittance matrix $\bm{Y}\equiv[Y_{ij}]_{i,j=1}^{n}\in\mathbb{C}^{n\times n}$. Specifically,
\begin{subequations}
\begin{align}
 P_{i} &= P_{i}^{\text{mech}} - P_{i}^{\text{load}} - |E_{i}|^{2} \Re\left(Y_{ii}\right) + \Re\left(E_i \cdot I_{i}^{*}\right),\label{defPi}\\
 \varphi_{ij} &= \begin{cases} -\arctan\left(\dfrac{\Re\left(Y_{ij}\right)}{\Im\left(Y_{ij}\right)}\right), & \text{if}\;i\neq j,\\
 0, & \text{otherwise},
 \end{cases}\\
 k_{ij} &= \begin{cases} |E_{i}| |E_{j}| \vert Y_{ij}\vert, & \text{if}\;i\neq j,\\
 0, & \text{otherwise},
 \end{cases}
\end{align}
\label{DefParam}
\end{subequations}
where $P_{i}^{\text{mech}}$ is the mechanical power input, $P_{i}^{\text{load}}$ is the real load, $E_{i}$ is the internal voltage, and $I_{i}\in\mathbb{C}$ is the current for generator $i$.

@@

%Suppose we denote the collection of generator buses in an \emph{unreduced} power system as $\mathcal{G}$, and the collection of non-generator buses as $\mathcal{N}$. 
Suppose that the \emph{unreduced} power network has $n$ generators and $m$ buses. Then the unreduced admittance matrix $\bm{Y}_{\rm{unreduced}}\in \mathbb{C}^{(n+m)\times(n+m)}$ can be partitioned as
\begin{align}
\bm{Y}_{\rm{unreduced}} = \begin{bmatrix} 
\begin{array}{c|c}
\bm{Y}_{\rm{bnd}} & \bm{Y}_{\rm{bnd{\text{-}}int}} \\
\hline\vspace*{-0.12in}\\
\bm{Y}_{\rm{bnd{\text{-}}int}}^{\top} & \bm{Y}_{\rm{int}}
\end{array}	
 \end{bmatrix},
\label{PartionedAdmittanceMatrix}	
\end{align} 
where $\bm{Y}_{\rm{bnd}}\in\mathbb{C}^{n\times n}$, $\bm{Y}_{\rm{int}}\in\mathbb{C}^{m\times m}$, $\bm{Y}_{\rm{bnd{\text{-}}int}}\in\mathbb{C}^{n\times m}$. The matrix \eqref{PartionedAdmittanceMatrix} relates the unreduced current vector $I_{\rm{unreduced}}\in\mathbb{C}^{n+m}$ with the unreduced voltage vector $E_{\rm{unreduced}}\in\mathbb{C}^{n+m}$ via the Kirchhoff equations
\begin{align}
I_{\rm{unreduced}} = \bm{Y}_{\rm{unreduced}} E_{\rm{unreduced}},
\label{DefUnreducedY}	
\end{align}
or equivalently, via its partitioned version associated with the interior and the boundary nodes:
\begin{align}
\begin{array}{c}
\begin{bmatrix}
I_{\rm{bnd}}\\
I_{\rm{int}} 	
\end{bmatrix}	
\end{array} = \bm{Y}_{\rm{unreduced}} \begin{array}{c}
\begin{bmatrix}
E_{\rm{bnd}}\\
E_{\rm{int}} 	
\end{bmatrix}	
\end{array}.
\label{PartitionedKirchoffEquation}	
\end{align}
%The reduced admittance matrix
%\begin{eqnarray}
%\bm{Y} &=& Y_{\mathcal{G}\mathcal{G}} - Y_{\mathcal{G}\mathcal{L}} Y_{\mathcal{L}\mathcal{L}}^{-1} Y_{\mathcal{L}\mathcal{G}} \\
%I_{red} &=& Y_{\mathcal{G}\mathcal{L}}Y_{\mathcal{L}\mathcal{L}}^{-1} I
%\end{eqnarray}

The network reduced admittance matrix $\bm{Y}\in\mathbb{C}^{n\times n}$ in \eqref{DefParam} is the Schur complement of \eqref{PartionedAdmittanceMatrix} w.r.t. the block $\bm{Y}_{\rm{int}}$, i.e.,
\[\bm{Y} = \bm{Y}_{\rm{unreduced}} / \bm{Y}_{\rm{int}} := \bm{Y}_{\rm{bnd}} - \bm{Y}_{\rm{bnd{\text{-}}int}} \bm{Y}_{\rm{int}}^{-1}\bm{Y}_{\rm{bnd{\text{-}}int}}^{\top}.\]

The network reduced current vector $I\in\mathbb{C}^{n}$ in \eqref{defPi} is obtained as
\[I  = \bm{Y}E_{\rm{bnd}} - I_{\rm{bnd}}.\]

@@

\begin{figure}[tb]
\centering
\includegraphics[width=0.9\linewidth]{ProxSchmatic.pdf}
\caption{\small{A schematic of approximating the flow of the joint PDF trajectory $\rho(t,\bm{x})$ via the sequence $\{\varrho_{k}\}_{k\in\mathbb{N}}$ generated by a variational recursion over $\mathcal{P}_{2}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right)$. The curve $\rho(t,\bm{x})$ solves a kinetic Fokker-Planck PDE initial value problem. While the joint PDFs are supported over the finite dimensional base manifold $\mathbb{T}^{n}\times\mathbb{R}^{n}$, the proximal updates $\{\varrho_{k}\}_{k\in\mathbb{N}}$ (shown as circled markers with no face-color) evolve over the infinite dimensional manifold $\mathcal{P}_{2}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right)$.}}
\vspace*{-0.1in}
\label{fig:ProxSchematic}
\end{figure}
